---
title: "project-lib"
author: "Collaborative class project"
date: "2026-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task Allocation

In a list, provide your name and which task you completed

## Reproducible Data Science

### The Importance of Reproducible Research: Main Steps

Here, you should explain why research needs to be reproducible and transparent, and provide a brief overview of the key steps involved.

### Open Science and FAIR Data Principles

Summarise the main principles of Open Science and explain what FAIR Data Principles stand for. Explain the main steps one should take to make their work reproducible.

### Resources and Platforms 

Provide brief descriptions and links to the resources that one could use to make their work reproducible.


## The Art (and Science) of Plotting

### Plot Types

Provide an overview of the main plot types (including visual examples) and explain which types of data are best suited for each kind of plot.

### Colour Palettes

Explain how to use colour palettes effectively, which types of palettes are best suited to different kinds of data, and what factors should be considered when selecting an appropriate colour scheme.

### Plotting Tools and Resources

Here, you should include links and brief descriptions of platforms and software for data visualization and colour palette selection, as well as useful R packages for plotting and styling graphics.


## Linguistic Database Repository

### Database Types

Summarise the main types of linguistic databases and describe the key features that distinguish them from one another.

### Choosing Database for your Research

Explain which types of linguistic databases are best suited to different major research directions in linguistics.

### Links and Descriptions

Provide links to several linguistic databases and briefly summarise the type of data each one contains.


## Handling messy data with *tidyverse*

### *Tidyverse* umbrella

Explain which packages are included in *tidyverse* and describe the purpose of each package and the types of tasks it is best suited for.

### *Dplyr* cheat sheet

Provide the list of most useful *dplyr* commands, explain what they do, and demonstrate how they work

### *Tidyr* cheat sheet

Provide the list of most useful *tidyr* commands, explain what they do, and demonstrate how they work

### *Ggplot2* cheat sheet

Provide the list of *ggplot2* commands and aesthetic parameters, explain what they do, and demonstrate which changes they make to a plot


## Compiling an R Markdown document

### Markdown Output Formats

Summarise the main output formats (not all available options) and explain what each is useful for.

### Markdown Text Formatting

Demonstrate how to format text, create lists, provide links, embed images, create footnotes

### Markdown Citations

Explain how to create cross-references within a document, generate a bibliography, and properly add citations.

### Formatting Tables 

#Explain how to create tables and captions, and describe the main table formatting options available in R Markdown.

There are different options of creating tables in an R Markdown file.

#### 1. Using 'Pipes' for simple tables:

```{r}
#Table: Table 1: Names, ages and animals

#| Name | Age | Animal |
#|------|-----|------|
#| Mirjam | 21  | Duck |
#| Sally | 25  | Frog   |
#| Dorina | 22  | Raccoon  |
```

Table: Table 1: Names, ages and animals

| Name | Age | Animal |
|------|-----|------|
| Mirjam | 21  | Duck |
| Sally | 25  | Frog   |
| Dorina | 22  | Raccoon  |

To add a caption to your table, you just need to write your desired caption on top of the table with a preceeding "Table:".

#### 2. Using 'knitr::kable()' function:

```{r}
library(knitr)

data <- data.frame(
    Name = c("Mirjam", "Sally", "Dorina"),
    Age = c(21.2202, 25.1703, 22.2802),
    Animal = c("Duck", "Frog", "Raccoon")
)

kable(data, caption = "Table 2: Names, ages and animals")
```
To add a caption, just define your desired caption in the "caption" variable.

#### 2.1 Formatting using 'knitr::kable()':
Usign 'knitr' you can format your data easily.
- Alignment: Aligning the columns left, center or right
```{r}
kable(data, align = c("l", "c", "r"))
```
- Number formatting: How many decimals should be displayed
```{r}
kable(data, digits = 2)
```

#### 3. Using 'table()' function to count frequencies:

```{r}
Animals <- c("Duck", "Duck", "Frog", "Raccoon", "Frog")

#table(animals)  <--- This works and produces a table, with that you just won't be able to add a caption
knitr::kable(table(Animals), caption = "Table 3: Frequency of animals")
```
To add a caption, you need to wrap the 'table()' function into the kable() function and define your 'caption' variable.
Using 'kable()' also makes the table a lot prettier than just the plain 'table()' table.

#### Further literature on RMD tables:
I found this interesting paper from the R-Project organisation themselves. It's a detailed paper on how to create different and more complex tables than just these three rather simple ones introduced.
If you're interested, go check it out: [Tables in R PDF URL](https://cran.r-project.org/web/packages/DescTools/vignettes/TablesInR.pdf)

### Code Cells

Summarise the main parameters used to create and format code chunks in R Markdown, and explain how these options affect the resulting code output and document rendering.
